---
title: "Section 4: Course Project - Hands-on Experience!"
author: "Robet Phavong"
date: "2024-09-27"
output: html_document
---


# Load necessary libraries
```{r}
library(shiny)
library(shinythemes)
library(ggplot2)
library(DT)
```


# Take a Look at the Project Data
## Background on data
### Approx, 50 observations = mining stocks
### MarketCap = market capitalization in million
### Symbol = abbreviation for comp. name (randomized)
### G1 = number of most valuable plots
### G2 = number of intermediate plots
### G3 = numver of least valuable plots
### Preferences of investors might differ
### Mining stock evaluation:
#### the investor specifies the importance of each grade(weight)
#### each grade is multiplied by its weight
#### figures are added together to get a score
#### evaluation is based on the scores
```{r}
# Load the data to be used for this project
mydata <- read.csv("course_proj_data.csv", header = T, sep = ";") 
```


# Project Description - Your Homework
### Data structure: sliders -> plot with brush -> table
### Three types of mining plots: 
#### Grade 1: most valuable
#### Grade 2: intermediate plots
#### Grade 3: least valuable


# Project Solution - User Interface
```{r}
server <- function(input, output, session) {}

ui <- navbarPage(theme = shinytheme("sandstone"), title = h3("The Mining Stock Scale"),
                 
                 tabPanel(("Adjust your Mining Stocks"),
                          wellPanel(
                            sliderInput(inputId = "w1",
                                        label = "Weight on Grade 1",
                                        value = 7, min = 0, max = 20),
                            sliderInput(inputId = "w2",
                                        label = "Weight on Grade 2",
                                        value = 2, min = 0, max = 20),
                            sliderInput(inputId = "w3",
                                        label = "Weight on Grade 3",
                                        value = 0.6, min = 0, max = 6, step = 0.2)
                            ),
                          plotOutput("scat", brush = "user_brush"),
                          DT::dataTableOutput("table"),
                          downloadButton(outputId = "mydownload", label = "Download Table")
                          ),
                 tabPanel("Documentation",
                          h4("Video documentation - Embedded from Youtube"),
                          tags$iframe(src="https://www.youtube.com/embed/vySGuusQI3Y"),
                          style="height:900px; width:900px"),
                 
                 tabPanel("Data Table with the underlying data",
                          DT::dataTableOutput("tableDT"))
                 
                 )

shinyApp(ui = ui, server = server)
```


# Project Solution - Server
```{r}
attach(mydata)

# Server section
server <- function(input, output, session) {
  
  ## Table for the Data table tab
  output$tableDT <- DT::renderDataTable(DT::datatable(mydata) %>%
                                          formatCurrency("MarketCap.in.M", "$", digits = 0) %>%
                                          formatStyle("Symbol", color = "grey") %>%
                                          formatStyle(c("G3", "G2", "G1"),
                                                      backgroundColor = "lightblue"))
  
  # Working on how the data will be calculated for the plot
  weighted.mydata = reactive(
    cbind(mydata, 
          points = input$w1 * `G1` + input$w2 * `G2` + input$w3 * `G3`) # be careful, use back quotes "``" and not single quots "''"
  )
  
  ## Working on the Plot
  output$scat = renderPlot({
    ggplot(weighted.mydata(), aes(points, MarketCap.in.M)) +
      geom_point() + geom_smooth(method = "lm") + 
      xlab("Your Calculated Score") + ylab("Market Capitalization in Million USD")
  })
  
  # What type of format to select for the data points of interest
  mydata.new = reactive({
    
    user_brush <- input$user_brush
    mysel <- brushedPoints(weighted.mydata(), user_brush)
    return(mysel)
    
  })
  
  # This table will be output when the user brushes onto a section of interest
  output$table = DT::renderDataTable(DT::datatable(mydata.new()))
  
  # Include a download tab of the data
  output$mydownload = downloadHandler(
    filename = "selected_miners.csv",
    content = function(file) {
      write.csv(mydata.new(), file)})
}
  

# User Interface section
ui <- navbarPage(theme = shinytheme("sandstone"), title = h3("The Mining Stock Scale"),
                 
                 tabPanel(
                   ("Adjust your Mining Stocks"),
                   wellPanel(
                     sliderInput(inputId = "w1",
                                 label = "Weight on Grade 1",
                                 value = 7, min = 0, max = 20),
                     sliderInput(inputId = "w2",
                                 label = "Weight on Grade 2",
                                 value = 2, min = 0, max = 20),
                     sliderInput(inputId = "w3",
                                 label = "Weight on Grade 3",
                                 value = 0.6, min = 0, max = 6, step = 0.2)
                     ),
                   plotOutput("scat", brush = "user_brush"),
                   DT::dataTableOutput("table"),
                   downloadButton(outputId = "mydownload", label = "Download Table")
                   ),
                 tabPanel("Documentation",
                          h4("Video documentation - Embedded from Youtube"),
                          tags$iframe(src="https://www.youtube.com/embed/vySGuusQI3Y"),
                          style="height:900px; width:900px"),
                 
                 tabPanel("Data Table with the underlying data",
                          DT::dataTableOutput("tableDT"))
                 
                 )


shinyApp(ui = ui, server = server)
```



## Copied script from professor for cross reference
```{r}
### Project Shiny

library(shiny)
library(DT)
library(ggplot2)
library(shinythemes)

mydata = read.csv("course_proj_data.csv", header = T, sep = ";")

attach(mydata)

server = function(input, output, session){
 
  # table for the Data table tab
 
  output$tableDT <- DT::renderDataTable(DT::datatable(mydata) %>%
                                          formatCurrency("MarketCap.in.M", "$", digits = 0) %>%
                                          formatStyle("Symbol", color = "grey") %>%
                                          formatStyle(c("G3", "G2", "G1"),
                                                      backgroundColor = "lightblue"))
 
 
  weighted.mydata = reactive(
    cbind(mydata,
          points = input$w1 * `G1` + input$w2 * `G2` + input$w3 * `G3`)
  )
 
 
  output$scat = renderPlot({
    ggplot(weighted.mydata(), aes(points, MarketCap.in.M)) +
    geom_point() + geom_smooth(method = "lm") +
    xlab("Your Calculated Score") + ylab("Market Capitalization in Million USD")
  })
 
 
  mydata.new = reactive({
    
    user_brush <- input$user_brush
    mysel <- brushedPoints(weighted.mydata(), user_brush)
    return(mysel)
    
  })
 
 
  output$table = DT::renderDataTable(DT::datatable(mydata.new()))
 
  output$mydownload = downloadHandler(
    filename = "selected_miners.csv",
    content = function(file) {
      write.csv(mydata.new(), file)})
 
 
}

ui = navbarPage(theme = shinytheme("sandstone"), title = h3("The Mining Stock Scale"),
                tabPanel(
  ("Adjust your Mining Stocks"),
 wellPanel(
  sliderInput(inputId = "w1",
              label = "Weight on Grade 1",
              value = 7, min = 0, max = 20),
  sliderInput(inputId = "w2",
              label = "Weight on Grade 2",
              value = 2, min = 0, max = 20),
  sliderInput(inputId = "w3",
              label = "Weight on Grade 3",
              value = 0.6, min = 0, max = 6, step = 0.2)
  ),
  plotOutput("scat", brush = "user_brush"),
  DT::dataTableOutput("table"),
  downloadButton(outputId = "mydownload", label = "Download Table")
),

tabPanel("Documentation",
         h4("Video documentation - Embedded from Youtube"),
         tags$iframe(style="height:700px; width:100%",
                  src="https://www.youtube.com/embed/vySGuusQI3Y")
         ),
         
tabPanel("Data Table with the underlying Data",
         DT::dataTableOutput("tableDT"))

)

shinyApp(ui = ui, server = server)
```


