---
title: 'Section3: Publishing and Advanced Concepts of Shiny'
author: "Robet Phavong"
date: "2024-09-27"
output: html_document
---


# Load necessary libraries
```{r}
library(shiny)
library(shinythemes)
library(ggplot2)
library(DT)
```


# Publishing and Sharing
## End user does not need R skills to use a Shiny app
## The coding part is made by the app developer
## Interactively through widgets
## R, Shiny and all required packages need to be installed on the server
## Setup your own server or use RStudio's server (shinyapps.io)
## Testing apps locally:
### Unsaved script: mark the whole app and run the code - ad hoc testing
### Saved script: no highlighting is needed, just run the code
#### Viewer pane, new window, Web browser (external)
### Works even offline!

## Easiest way: Shinyapps.io
### RStudio's server service
### Upload your code and share
### Different plans are available: free plan or monthly subscription
### R, Shiny and other add-ons are already installed

## Alternatively, setup your own server
### It needs to run R, Shiny and all required packages
### Linux based server
### Free server software from RStudio: RServer Basic
### Pro version comes with a price
### Setup and maintenance takes effort and requires a skillsed programmer

## Script naming convention
### Proper script naming convention is required if app is deployed onto a server
## Double script - old method
### server.R - server part of the code
### ui-R - user interface part of the code
### Appropriate for long scripts

## Mono script - common method
### app.R - contains both parts of the code

## R Server recognizes scripts only as app.R or server.R and ui.R
## Each component of the app must be in one folder
## Local scripts can be named whatever you want


# Shinyapps.io
## Option 1 (hard way): Own server - Linux server with Shiny Server software OR Linux server with Shiny Server Pro (paid with authentication)

## Option 2: Shinyapps.io - service provided by RStudio. Host and share your apps

# Preparing shinyapps.io account
## 1. Create Shinyapps.io account using the shinyapps.io website
## 2. install.packages('rsconnect')
## 3. Authorize account: copy the account token and secret
```{r}
rsconnect::setAccountInfo(name='rphavong',
			  token='1FE2E12E29F096FF62D2FC812D021C8B',
			  secret='<SECRET>')
```


## 4. Deploy your application to shinyapps.io
```{r}
library(rsconnect)
    rsconnect::deployApp('path/to/your/app')
```


# Integrating Tables with DT
## Tables can be a part of or the whole Shiny app
## Well implemented in Shiny
## Shiny function: renderDataTable() -> sets up a data table
## DT package: renderDataTable()
### Make sure you use the right function: DT::renderDataTable()
### DT generates interactive, visually appealing data tables
### Install the package DT once and activate in each session
### DT is a rather new package (therefore not much documentation)
### DT integrates JavaScript library DataTables into R
## Documentation: JavaScript's DataTables library: datatables.net or DT packageL rstudio.github.io/dt
```{r}
## Basic approach
server <- function(input, output, session) {
  
  output$tableDT <- DT::renderDataTable(diamonds[1:1000,],
                                        options = list(paging=F),
                                        rownames=F,
                                        filter = "top")
}

ui <- fluidPage(
  DT::dataTableOutput("tableDT")
)

shinyApp(ui = ui, server = server)
```


```{r}
## Adding features to the table within the output section

server <- function(input, output, session) {
  
  output$tableDT <- DT::renderDataTable(datatable(diamonds[1:1000,],
                                                  options = list(paging=F),
                                                  rownames=F,
                                                  filter = "top") %>%
                                          formatCurrency("price", "$") %>%
                                          formatStyle("price", color = "green") %>%
                                          formatStyle("cut",
                                                      transform = "rotateX(20deg) rotateY(5deg) rotateZ(5deg)",
                                                      backgroundColor = styleEqual(
                                                        unique(diamonds$cut), c("salmon", "lightblue", "grey", "lightgreen", "lightpink"))))
}

ui <- fluidPage(
  DT::dataTableOutput("tableDT")
)

shinyApp(ui = ui, server = server)
```


# Add Brush and Click Selection tools to your App
## User Input
### UI input widgets
### 1. Data input through input widget (user end)
### 2. Calculations (server end)
### 3. Calculated results/Output (user end)

## Alternative types of user input:
### Click, double click, hover, brush movement, etc
### They work with R base plots and ggplot2 plots
### Our Example: 
#### Graph sensitive to a brush maneuver 
#### User action click
```{r}
## Advanced App - Brush Feature

server <- function(input, output, session) {
  
  # Include which R packages need to be loaded 
  library(ggplot2) # for the diamonds dataset, and ggplot features
  library(DT) # for the dataTableOutput function
  
  # Include which type of plot to use and what to include in the x and y axis
  output$plot <- renderPlot({
    ggplot(diamonds, aes(price, carat)) + geom_point()
  })
  
  # Make graph reactive by user by creating a function
  diam <- reactive({
    user_brush <- input$user_brush
    sel <- brushedPoints(diamonds, user_brush) # note here we are including our dataframe of interest 'diamonds' and the user_brush to include by user
    return(sel)
  })
  
  output$table <- DT::renderDataTable(DT::datatable(diam()))
}

ui <- fluidPage(
  h1("Using the brush feature to select specific observations"),
  plotOutput("plot", brush = "user_brush"),
  dataTableOutput("table")
)

shinyApp(ui = ui, server = server)
```


## Similar to code above, but suing the click feature
```{r}
## Advanced App - Click Feature

server <- function(input, output, session) {
  
  # Include which R packages need to be loaded 
  library(ggplot2) # for the diamonds dataset, and ggplot features
  library(DT) # for the dataTableOutput function
  
  # Include which type of plot to use and what to include in the x and y axis
  output$plot <- renderPlot({
    ggplot(diamonds, aes(price, carat)) + geom_point()
  })
  
  # Make graph reactive by user by creating a function
  diam <- reactive({
    
    user_click <- input$user_click
    sel <- nearPoints(diamonds, user_click, threshold = 10, maxpoints = 5) # note here we are including our dataframe of interest 'diamonds' and the user_click to include by user
    return(sel)
  })
  
  output$table <- DT::renderDataTable(DT::datatable(diam()))
}

ui <- fluidPage(
  h1("Using the click feature to select specific observations"),
  plotOutput("plot", click = "user_click"),
  dataTableOutput("table")
)

shinyApp(ui = ui, server = server)
```


# Downloadable Tables
```{r}
## Advanced Plot with CSV export of data set

server <- function(input, output, session) {
  
  # Include which R packages need to be loaded 
  library(ggplot2) # for the diamonds dataset, and ggplot features
  library(DT) # for the dataTableOutput function
  
  # Include which type of plot to use and what to include in the x and y axis
  output$plot <- renderPlot({
    ggplot(diamonds, aes(price, carat)) + geom_point()
  })
  
  # Make graph reactive by user by creating a function
  diam <- reactive({
    user_brush <- input$user_brush
    sel <- brushedPoints(diamonds, user_brush) # note here we are including our dataframe of interest 'diamonds' and the user_brush to include by user
    return(sel)
  })
  
  output$table <- DT::renderDataTable(DT::datatable(diam()))
  
  # Include data to be easily accessible to user to be downloadable as a CSV file
  output$mydownload <- downloadHandler(
    filename = "plotextract.csv",
    content = function(file) {
      write.csv(diam(), file)})
}

ui <- fluidPage(
  h1("Using the brush feature to select specific observations"),
  plotOutput("plot", brush = "user_brush"),
  dataTableOutput("table"), 
  downloadButton(outputId = "mydownload", label = "Download Table") # Include a download button/tab for user to easily click to download data in CSV format
)

shinyApp(ui = ui, server = server)
```


# Integrating all Sorts of Media
## Media integration
### Video, audio, image, iframe, text, PDF
### Most of them can be integrated via tags
#### ExceptionL text has its own function - includeText()
#### Tags: tags$video, tags$img, tags$iframe, tags$audio
#### Local data
#### Framed data (online)
### Shiny apps can be integrated into websites via iframes
```{r}
library(shiny)

server <- function(input, output) {}

ui <- navbarPage("Integration of different media types",
                 
                 tabPanel("Image sourced localled", 
                          tags$img(src = "logo.png", width = "100px", height = "100px")),
                 
                 tabPanel("Video sourced locally",
                          tags$video(src = "comist.mp4", type = "video/mp4", controls = T,
                                     width = "900px", height = "800px")),
                 
                 tabPanel("Pdf sourced online, Iframe",
                          tags$iframe(style="height:600px; width:100%; scrolling=yes",
                                      src="https://cran.r-project.org/web/packages/shiny/shiny.pdf")),
                 
                 # For this tab you'll need to create a txt file using the indicated file name as templated below "mytxt.txt"
                 # tabPanel("Text as .txt",
                 #          includeText("mytxt.txt"))

)

shinyApp(ui = ui, server = server)

```


# Tools for your own Apps
## Can look up Shiny by RStudio Gallery for all R scripts to use the featured visualizations already created. Can use the code to then modify at your own discretion. 
